On-the-fly Algorithms for GKAT Equivalences \thanks{Identify applicable funding agency here. If none, delete this.}

### 2025-01-19 main.tex
revise evaluation section

\section{Implementation}
We implement our symbolic on-the-fly bisimulation algorithm~(\Cref{alg:symb-bisim}) in Rust along with derivative based~(\Cref{fig:derivatives-rules}) and Thompson's construction based~(\Cref{tab:symb-Thompson-construction}) algorithms for constructing symbolic GKAT coalgebras. The source code and benchmark suite is freely available in our repository TODO.

@@

\subsection{Performance}\label{sec:performance-implementation}
Our experiments are performed on a laptop with an Apple M4 Pro CPU and 24 GB RAM. \Cref{tab:benchmark-time,tab:benchmark-memory} present the results of equivalence checking based on derivatives (DV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015} (SK). Each benchmark consists of 50 expression pairs with properties described by the benchmark's name. For example, expressions in E250B5P10RD have around 250 primitive actions in total (E250), a maximum boolean expression size of 5 (B5), 10 possible unique primitive tests (P10) and are completely random (RD). Benchmarks with the suffix EQ have expression pairs that are known to be equivalent. The time columns measure the \textit{total time} used (in seconds) for checking all expression pairs in the benchmarks. The mem columns measure the \textit{peak memory} used (in MB).

It is clear from these results that our algorithms (DRV and TC) can efficiently check even large programs in just a few seconds.

\begin{table}
\footnotesize
\centering
\setlength{\tabcolsep}{3pt}
\begin{tabular}{l r r r r r}
    Benchmark & DV$_{\textsf{BDD}}$ & TC$_{\textsf{BDD}}$ & DV$_{\textsf{SAT}}$ & TC$_{\textsf{SAT}}$ & SK \\
    \hline
    e250b5p10rd    &   0.19 &   0.18 & 0.20 & 0.23 &    5.82 \\
    e250b5p10eq    &   0.21 &   0.18 & 0.18 & 0.29 &    2.83 \\
    e500b5p50rd    &   0.23 &   0.22 & 0.27 & 0.39 &   37.28 \\
    e500b5p50eq    &   0.26 &   0.21 & 0.28 & 0.57 &   14.06 \\
    e1000b10p100rd &   0.31 &   0.37 & 0.48 & 0.85 & timeout \\
    e1000b10p100eq &   0.37 &   0.28 & 0.44 & 1.04 &   77.83 \\
    e2000b20p200rd &   1.50 &   2.75 & 0.99 & 2.01 & timeout \\
    e2000b20p200eq &   3.34 &   3.03 & 0.67 & 3.14 & timeout \\
    e3000b30p200rd &   1.75 &  26.38 & 1.91 & 4.46 & timeout \\
    e3000b30p200eq &  17.86 &  22.59 & 1.09 & 5.25 & timeout \\
    degenerate     & 220.53 & 232.63 & 0.42 & 0.93 & timeout
\end{tabular}
\caption{Benchmark Time Usage}\label{tab:benchmark-time}
\end{table}

\begin{table}
\footnotesize
\centering
\setlength{\tabcolsep}{3pt}
\begin{tabular}{l r r r r r}
    Benchmark & DV$_{\textsf{BDD}}$ & TC$_{\textsf{BDD}}$ & DV$_{\textsf{SAT}}$ & TC$_{\textsf{SAT}}$ & SK \\
    \hline
    e250b5p10rd    &  15.36 &   14.76 &  6.81 &  7.02 &  114.06 \\
    e250b5p10eq    &  15.56 &   14.95 &  7.06 &  7.25 &  100.48 \\
    e500b5p50rd    &  16.26 &   15.49 &  7.02 &  7.25 &  524.89 \\
    e500b5p50eq    &  17.97 &   15.54 &  7.07 &  7.27 &  546.91 \\
    e1000b10p100rd &  18.36 &   21.41 &  8.22 &  7.99 & timeout \\
    e1000b10p100eq &  20.41 &   17.66 &  8.69 &  7.63 & 5822.66 \\
    e2000b20p200rd & 239.71 &  283.45 & 13.33 & 11.92 & timeout \\
    e2000b20p200eq & 107.92 &  102.46 & 13.20 & 14.02 & timeout \\
    e3000b30p200rd & 112.59 & 1235.46 & 21.44 & 20.10 & timeout \\
    e3000b30p200eq & 245.92 &  228.85 & 21.41 & 18.15 & timeout \\
    degenerate     & 631.47 & 1229.30 & 19.62 & 18.43 & timeout
\end{tabular}
\caption{Benchmark Memory Usage}\label{tab:benchmark-memory}
\end{table}

>>>>

\section{Implementation and Evaluation}
We implement our symbolic on-the-fly bisimulation algorithm~(\Cref{alg:symb-bisim}) in Rust along with derivative based~(\Cref{fig:derivatives-rules}) and Thompson's construction based~(\Cref{tab:symb-Thompson-construction}) algorithms for constructing symbolic GKAT coalgebras. The source code and benchmark suite is freely available in our repository TODO.

@@

\subsection{Evaluation}\label{sec:performance-implementation}
Due to the fact that our symbolic algorithms are generic regarding the choice of solvers for boolean satisfiability, we compare their performance using BDD to using a SAT solver (MiniSat~\cite{een_MINISAT_2004}). Our experiments are performed on a laptop with an Apple M4 Pro CPU and 24 GB RAM. \Cref{tab:benchmark-time,tab:benchmark-memory} present the experimental results of equivalence checking based on derivatives (DV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015} (SK). Note that SymKAT uses BDD internally for checking boolean satisfiability. Each benchmark consists of 50 expression pairs with properties described by the benchmark's name. For example, expressions in `e250b5p10eq' have approximately 250 primitive actions in total (e250), a maximum boolean expression size of 5 (b5), 10 possible unique primitive tests (p10) and are known to be equivalent (eq). Benchmarks with the suffix `ne' have expression pairs that are known to be non-equivalent.

\begin{table}[H]
\footnotesize
\centering
\setlength{\tabcolsep}{3pt}
\begin{tabular}{l r r r r r}
    Benchmark & DV$_{\textsf{BDD}}$ & TC$_{\textsf{BDD}}$ & DV$_{\textsf{SAT}}$ & TC$_{\textsf{SAT}}$ & SK \\
    \hline
    e250b5p10ne    &   0.19 &   0.18 & 0.20 & 0.23 &    5.82 \\
    e250b5p10eq    &   0.21 &   0.18 & 0.18 & 0.29 &    2.83 \\
    e500b5p50ne    &   0.23 &   0.22 & 0.27 & 0.39 &   37.28 \\
    e500b5p50eq    &   0.26 &   0.21 & 0.28 & 0.57 &   14.06 \\
    e1000b10p100ne &   0.31 &   0.37 & 0.48 & 0.85 & timeout \\
    e1000b10p100eq &   0.37 &   0.28 & 0.44 & 1.04 &   77.83 \\
    e2000b20p200ne &   1.50 &   2.75 & 0.99 & 2.01 & timeout \\
    e2000b20p200eq &   3.34 &   3.03 & 0.67 & 3.14 & timeout \\
    e3000b30p200ne &   1.75 &  26.38 & 1.91 & 4.46 & timeout \\
    e3000b30p200eq &  17.86 &  22.59 & 1.09 & 5.25 & timeout \\
    degenerate     & 220.53 & 232.63 & 0.42 & 0.93 & timeout
\end{tabular}
\caption{Total Time Usage}\label{tab:benchmark-time}
\end{table}

\begin{table}[H]
\footnotesize
\centering
\setlength{\tabcolsep}{3pt}
\begin{tabular}{l r r r r r}
    Benchmark & DV$_{\textsf{BDD}}$ & TC$_{\textsf{BDD}}$ & DV$_{\textsf{SAT}}$ & TC$_{\textsf{SAT}}$ & SK \\
    \hline
    e250b5p10ne    &  15.36 &   14.76 &  6.81 &  7.02 &  114.06 \\
    e250b5p10eq    &  15.56 &   14.95 &  7.06 &  7.25 &  100.48 \\
    e500b5p50ne    &  16.26 &   15.49 &  7.02 &  7.25 &  524.89 \\
    e500b5p50eq    &  17.97 &   15.54 &  7.07 &  7.27 &  546.91 \\
    e1000b10p100ne &  18.36 &   21.41 &  8.22 &  7.99 & timeout \\
    e1000b10p100eq &  20.41 &   17.66 &  8.69 &  7.63 & 5822.66 \\
    e2000b20p200ne & 239.71 &  283.45 & 13.33 & 11.92 & timeout \\
    e2000b20p200eq & 107.92 &  102.46 & 13.20 & 14.02 & timeout \\
    e3000b30p200ne & 112.59 & 1235.46 & 21.44 & 20.10 & timeout \\
    e3000b30p200eq & 245.92 &  228.85 & 21.41 & 18.15 & timeout \\
    degenerate     & 631.47 & 1229.30 & 19.62 & 18.43 & timeout
\end{tabular}
\caption{Peak Memory Usage}\label{tab:benchmark-memory}
\end{table}

We observe that all variants of DV and TC perform better than SymKAT in terms of time and memory usage. When comparing BDD to SAT, notice that BDD is marginally faster than SAT for smaller benchmarks, however, its time and memory usage degrades for larger ones. On the other hand, the performance of SAT is stable for all benchmarks. Furthermore, we have encountered BDD-degenerative expressions (the `degenerate' benchmark) which greatly reduce BDD performance but can be efficiently solved by SAT. We did not encounter any SAT-degenerative expressions where BDD performed significantly better than SAT.
